---
title: "Life Expectancy for Political Candidates"
author: "Umaira"
format: html
execute: 
  echo: false
---


```{r}
#| message: false
#| warning: false
library(tidyverse)
library(primer.data)
library(tidymodels)
library(broom)
library(gt)
```

Using data from over 1,000 candidates for governor in the United States since 1945, we aim to develop a model that estimates how age at election may relate to post-election longevity. However, since our data is limited to gubernatorial candidates, the model’s predictions may not generalize well to individuals running for other political offices or in different contexts. We model longevity after election, measured in years lived following a gubernatorial campaign, as a linear function of the candidate’s sex and their age at the time of election.

$$Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \cdots + \beta_n X_n + \epsilon$$

with $\epsilon \sim \mathcal{N}(0, \sigma^2)$.


$$
\hat{Y}_i = 40 - 0.5 \cdot \text{age}_i - 2.3 \cdot \text{won}_i + 0.04 \cdot \text{year}_i
$$

```{r}
#| cache: true
x <- governors |>
  filter(year >= 1945)
```

```{r}
#| cache: true
fit_years <- linear_reg(engine = "lm") |>
  fit(lived_after ~ sex + election_age, data = x)

```

```{r}
#| cache: true
tidy(fit_years, conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  gt() |>
  tab_header(
    title = "Model Estimates for Longevity After Election",
    subtitle = "Based on sex and age at election"
  ) |>
  fmt_number(
    columns = c(estimate, conf.low, conf.high),
    decimals = 2
  ) |>
  cols_label(
    term = "Variable",
    estimate = "Estimate",
    conf.low = "Lower 95% CI",
    conf.high = "Upper 95% CI"
  ) |>
  tab_source_note(
    source_note = "Source: U.S. Governors dataset (since 1945)"
  )


```
